Understanding models for spatial structure in small-area estimation

### 2023-05-30 src/docs_paper/paper.Rmd
Update paper

---
title: Understanding models for spatial structure in small-area estimation
author:
  - name: Adam Howes
    email: ath19@ic.ac.uk
    affiliation: A
    corresponding: ath19@ic.ac.uk
  - name: Jeffrey W. Eaton
    affiliation: B
  - name: Seth R. Flaxman
    affiliation: C
address:
  - code: A
    address: Department of Mathematics, Imperial College London
  - code: B
    address: MRC Centre for Global Infectious Disease Analysis, Imperial College London
  - code: C
    address: Department of Computer Science, University of Oxford
abstract: |
  |  **Background:** Spatial correlation between small-areas is typically accounted for within a generalised linear mixed model setting using spatial random effects. However, for some geometries, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic and unsatisfying spatial assumptions. Furthermore, they do not take into account that across many settings areal data is generated by aggregating continuous point data. Defining spatial random effects which instead correspond to aggregated Gaussian processes gives a more intuitively convincing correlation structure, and allows tools from kernel methods to be used in models for areal data.
  |
  |  **Methods:** In a simulation study, we used a proper scoring rule to evaluate the performance of different spatial random effect specifications. We performed these simulations under model misspecification and across a range of vignette and realistic geometries with different regularities. We also compared model performance in estimating district-level HIV prevalence from PHIA household survey data across four countries in sub-Saharan Africa, using cross-validation and spatial cross-validation.
  |
  |  **Results:**
  |
  |  **Conclusions:**

@@

bibliography: citations.bib
output:
  rticles::elsevier_article:
    number_sections: true
    includes:
      in_header: preamble.tex
# plos.csl can be download and used locally
csl: http://www.zotero.org/styles/plos
---

@@

Spatial random effects are frequently used to model spatial variation in areal data [@haining2003spatial; @cramb2018investigation].
Gaussian Markov random fields (GMRFs) [@rue2005gaussian], which combine a Gaussian distribution with Markov conditional independence assumptions between areas, are the most common class of models used to specify spatial random effects.
Using a GMRF typically reflects the assumption that observations made in areas close together are correlated, and more distant relationships may be ignored.
The simplest GMRF model is arguably that of @besag1991bayesian, where information is borrowed equally from each adjacent area.
This model requires minimal additional modelling choices and is widely implemented, including as a part of the `R-INLA` package [@rue2017bayesian], popular within spatial statistics.
Use of this model is widespread, including in agriculture [@oliver2015soil], ecology [@saracco2010modeling], epidemiology [@lawson2013statistical], image analysis [@schmid2006bayesian], neuroscience [@gossl2001bayesian] and public health [@dwyer2015drinking].
However, for irregular geometries like the administrative divisions of a country, the assumptions made by the Besag model about space are unrealistic.
In this work, we set out to test the hypothesis that incorporating spatial structure more faithfully to the specific geometry would improve the performance of small-area estimation models, and in doing so provide practical recommendations and further understanding of models for spatial structure.
<!-- @gutreuter2019improving use the SAR model, which is still adjacency based but is not exactly the Besag model--restructure introduction to be more general and include introduction of other adjacency based models (which all have the same criticisms)? -->

Our motivating application is the small-area estimation of HIV epidemic indicators in sub-Saharan Africa.
Estimates are required to help plan, implement, and evaluate the success of programmes, ensuring that available resources are most effectively used to respond to the epidemic [@hallett2016evaluation; @cuadros2017mapping].
Household surveys provide nationally represenative data about the general population, but are expensive to carry out and as such have small sample sizes at a district level.
Although auxiliary covariates can be used to aid estimation, data on the covariates most strongly associated with HIV, such as sexual risk behaviour or the prevalence of male circumcision, often face similar measurement difficulties as the HIV indicators themselves.
Models including covariates have only been found to modestly improve predictions \citep[Supplementary Figure 20]{dwyer2019mapping}.
This stands in contrast to other infectious diseases like Malaria where transmission is driven by more predictive and easily-measurable environmental factors [@weiss2015re].
These circumstances foreground the importance of sharing information between districts via spatial smoothing in HIV mapping.

Within the GRMF framework, attempts to more accurately reflect irregular spatial structure have defined weights specifying the extent to which neighbours are related.
@duncan2017spatial compared seventeen methods for specifying these weights, but surprisingly did not find any which outperform the Besag model, though this conclusion was specific to the often-analysed Scottish lip cancer example, and based on the deviance information criteria (DIC) which is recommended against by @vehtari2017practical.
Another approach is to take into account that areal data is typically produced by aggregating higher resolution data.
The problem is then a particular instance of the broader challenge of inference about a variable at a different resolution to that it was observed at, known in geostatistics as the change-of-support problem [@gelfand2001change], which dates to foundational work on block kriging [@krige1951statistical; @matheron1976forecasting].
Ideas from change-of-support modelling are applied by @kelsall2002modeling to consider areal data as coming about by an aggregation of a continuously-indexed Gaussian process, resulting in a covariance structure between two areas given by the average covariance between two points chosen randomly from those areas.
This type of model is particularly natural in the log-Gaussian Cox process modelling framework [@li2012log; @diggle2013spatial; @taylor2015bayesian; @johnson2019spatially].
<!-- `lgcp` [@taylor2015bayesian] and `SDALGCP` [@johnson2019spatially] packages. -->
Inference for aggregated data has recently been advanced by @wilson2018pointless who consider the SPDE approach of @lindgren2011explicit, using `R-INLA`, and an empirical Bayes approach, using `TMB` [@kristensen2015tmb], and made available as a part of the `disaggregation` package [@nandi2020disaggregation].

@@

@@

Let $\Sc \subset \mathbb{R}^2$ be the study region, and the disjoint areas $\{A_i\}_{i = 1}^n$ be a spatial partition of $\Sc$.
Areal data are a type of spatial data where observations $y_i$ are associated to areas $A_i$.
Examples include the colour of a pixel, the minimum wage in a state, and the number of disease cases in a region.
Point data are another type of spatial data where instead observations $y(s)$ can be made at any location $s \in \Sc$ of a spatially-continuous stochastic process.

Areal data is often aptly conceptualised as arising from aggregation of point data, such that $y_i = \int_{A_i} y(s) \text{d}s$.
Exceptions include policies determined at an administrative level, such as the minimum wage or disease control measures, which have a genuinely discrete spatial structure.
<!-- Note that, for data of this sort, there is less clear justification for using spatial smoothing. -->
<!-- Realistically, data relate to many covariates, some number of which are likely have discrete spatial structure and others of which continuous spatial structure, such that the result is often an intermediate. -->
<!-- It may be possible to model this explicitly, though the result might look similar to what is achieved using e.g. BYM2. -->
If the areal data we observe are indeed an aggregation of point data, why not use the higher resolution point data in our models rather than areal data?
Explanations include that (a) the underlying point data are unobserved, (b) although point data are collected, they may not be accessible due to privacy constraints, administrative practicality or storage capacity, and (c) the point data are available, but we decide not to use them in the model.
In the final case (c), researchers may prefer to use area-level models as models for point data are typically more complex, require more data, are less immediately applicable to area-level policy making, and have higher computational burden.
Furthermore, seemly point data may actual be areal, such as data observed at polling stations or health facilities which individuals from the surrounding area travel to.
As such, choosing to model the underlying data generating mechanism as either areal or point is often a matter of pragmatism.

@@

Bayesian hierarchical models are an attractive framework for small-area estimation, whereby areal data $y$ corresponding to the areas $\{A_i\}_{i = 1}^n$, are modelled using a three-layer structure [@berliner1996hierarchical; @cressie2015statistics; @rao2015small] given by
\begin{alignat}{2}
&\text{(Observations)}  &        y_i &\sim p(y_i \, | \, g^{-1}(x_i), \theta), \quad i = 1, \ldots, n, \label{eq:data} \\
&\text{(Latent field)}  &        x &\sim p(x \, | \, \theta), \label{eq:process} \\
&\text{(Parameters)}    & \qquad \theta &\sim p(\theta), \label{eq:parameters}
\end{alignat}
where $x$ is the $n$-dimensional latent field, $\theta$ are the $m$-dimensional parameters, $m < n$, and $g$ is a link function.
Spatial structure is modelled at the level of the latent field, and the observation model, typically either binomial or Poisson, is conditionally independent and identically distributed $p(y \, | \, x, \theta) = \prod_{i = 1}^n p(y_i \, | \, x_i, \theta)$.
As small-area estimation models are descriptive rather than mechanistic, it is natural [@best2005comparison] to use a flexible and computationally efficient Gaussian distribution $x \sim \mathcal{N}(\mu, \Sigma)$ for the latent field.
Three-layer models with Gaussian latent fields comprise a wide array of popular models, including generalised linear mixed models (GLMMs), and have been collectively studied under the title latent Gaussian models (LGMs) [@rue2009approximate].

@@

One way to encode spatial structure between areas is by using a symmetric relation, where $i \sim j$ if the areas $A_i$ and $A_j$ are adjacent or neighbouring.
Adjacency is often defined by a shared border, though other choices, such as inclusion of second or higher-degree neighbours [@paciorek2013spatial], are also possible.
The Besag model [@besag1991bayesian] is a type of improper conditional auto-regressive (ICAR) model where the full conditional distribution of the $i$th spatial random effect is given by
\begin{equation}
    \phi_i \, | \, \phi_{-i} \sim \N \left(\frac{1}{n_{\delta i}} \sum_{j: j \sim i} \phi_j, \frac{1}{n_{\delta i}\tau_\phi}\right), \label{eq:besag}
\end{equation}
where $\delta i$ is the set of neighbours of $A_i$ with cardinality $n_{\delta i} = |\delta i|$ and $\phi_{-i} = (\phi_1, \ldots, \phi_{i - 1}, \phi_{i + 1}, \ldots, \phi_n)^\top$ is the vector of spatial random effects with the $i$th entry removed.
The conditional mean of the random effect $\phi_i$ is the average of its neighbours $\{\phi_j\}_{j \sim i}$ and the precision $n_{\delta i}\tau_\phi$ is proportional to the number of neighbours $n_{\delta i}$.
By Brook's lemma \cite[Chapter 2]{rue2005gaussian} the set of full conditionals of the Besag model are equivalent to the Gaussian Markov random field (GMRF) given by
\begin{equation}
    \phi \sim \mathcal{N}(0, \tau_\phi^{-1} R^{-}), \label{eq:gmrf}
\end{equation}
where $R^{-}$ is the generalised inverse of the rank-deficient structure matrix $R$, so-called because it defines the \textit{structure} of the precision matrix, with entries
\begin{equation}
    R_{ij} =
    \begin{cases}
        n_{\delta i}, & i = j \\
        -1, & i \sim j \\
        0, & \text{otherwise.}
    \end{cases}
\end{equation}
The Markov property arises due to the conditional independence structure $p(\phi_i \, | \, \phi_{-i}) = p(\phi_i \, | \, \phi_{\delta i})$ whereby each area only depends on its neighbours.
This is reflected in the sparsity of $R$ whereby $\phi_i \perp \phi_j \, | \, \phi_{-ij}$ if and only if $R_{ij} = 0$.
The structure matrix $R$ may also be expressed as the Laplacian of the adjacency graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ with vertices $v \in \mathcal{V}$ corresponding to each area and edges $e \in \mathcal{E}$ between vertices $i$ and $j$ when $i \sim j$.
Figure \ref{fig:geometry-graph-zwe} shows the adjacency graph for the districts of Zimbabwe alongside the geometry.
Shortly, in Section \ref{sec:concerns} we discuss the appropriateness of using the graph rather than the complete geometry.

@@

Section \ref{sec:ab} reviewed ways to construct spatial random effect precision matrices using an adjacency relation.
Another approach is to define the covariance matrix using an areal kernel function $K: \mathcal{P}(\Sc) \times \mathcal{P}(\Sc) \to \mathbb{R}$ which gives a measure of similarity between two areas, where $\mathcal{P}$ denotes the power set and $\mathcal{P}(\Sc)$ is the space of subsets of the study region.
If $K$ is positive semi-definite, then we define areal kernel spatial random effects by
\begin{equation}
    \phi \sim \mathcal{N}(0, \frac{1}{\tau_\phi} K), \label{eq:arealkernel}
\end{equation}
where the $n \times n$ Gram matrix $K$ with entries $K_{ij} = K(A_i, A_j)$ is a valid covariance matrix.
Here, we place $\tau_\phi$ outside of the Gram matrix, analogous to the relation of the precision and structure matrices.

Most well-known spatial process models define the correlation structure between a pair of points using a kernel $k: \mathcal{S} \times \mathcal{S} \to \mathbb{R}$.
We use a simple method to construct $K$ from $k$, namely averaging the kernel $k$ computed on some collection of points from within each area.
As such, $K$ is described as a kernel on sets in the multiple instance learning literature [@gartner2002multi].
Although it may be possible to learn the method for combining kernels from data [@gonen2011multiple], we do believe that this is feasible in the small-area estimation setting, and instead hope to construct areal kernels which represent our prior beliefs about the spatial process.

@@

The simplest approach is to use a single point, especially one which is representative of the area, a natural choice being the centroid $c_i \in A_i$ given by the arithmetic mean of the latitude and longitude.
This results in the centroid kernel
\begin{equation}
    K(A_i, A_j) = k(c_i, c_j).
\end{equation}
<!-- More broadly, given a weighting distribution $W_i$ with probability density proportional to $w_i(s)$ over $A_i$, weighted centroids could be used in the above equation. -->
The centroid kernel has been used by @wakefield1999spatial in an environmental epidemiology setting and @best2005comparison in a model comparison study, both choosing the exponential kernel $k(s_i, s_j) = \exp \left( - |s_i - s_j| / l \right)$; as well as more recently by @teh2021efficient as a part of a spatio-temporal Gaussian process model to infer the reproduction number of COVID-19.
@wakefield1999spatial fix the length-scale $l$ using a semi-variogram and @best2005comparison place a prior on the inverse length-scale.
\citet[Section 3]{best2005comparison} simulated data representing heterogeneous exposure to air pollution, including elevated rates of exposure near two hypothetical point source locations, and found that the centroid kernel tended to over-smooth the high-risk areas, though it is unsuprising that a stationary covariance would struggle to recover non-stationary structure.

@@

Rather than attempting to choose a single representative point, an alternative is to represent the whole area by integrating the kernel over the areas of interest.
This results in the integrated kernel
\begin{equation}
K(A_i, A_j) = \frac{1}{|A_i||A_j|} \int_{A_i} \int_{A_j} k(s, s') \text{d} s \text{d} s', \label{eq:ik1}
\end{equation}
which is equivalent to the covariance structure obtained by aggregating a spatially continuous Gaussian process with kernel $k$ over the areal partition.
Models of this sort have been studied in the machine learning literature under the name aggregated Gaussian processes [@law2018variational; @tanaka2019spatially; @yousefi2019multi; @hamelijnck2019multi; @2021arXiv210501460Z].
<!-- % For example, @law2018variational develop a variational Bayesian inference method for aggregated observation models and @tanaka2019spatially consider aggregation over a linear combination of latent Gaussian processes, each of which is associated to data corresponding to a different partition of the study region. -->
Unlike for the centroid kernel where $K_{ii} = 1$ for all $i$, the marginal variance of the $i$th spatial random effect $K_{ii} = K(A_i, A_i)$ varies depending on the area: becoming smaller for more compact areas and larger for areas which are of greater extent or more spread out.

The log-Gaussian Cox Process framework [@diggle2013spatial] also naturally arrives at the integrated kernel formulation.
A Cox process is an inhomogeneous Poisson process with a continuous stochastic intensity function $\{ x(s), s \in \Sc \}$ such that conditional on the realisation of $x(s)$ the number of points in any area $A_i$ follows a Poisson distribution.
The rate parameter of this Poisson distribution is explicitly aggregated as follows
\begin{equation}
y_i \, | \, x(s) \sim \text{Poisson} \left(\int_{s \in A_i} x(s) \text{d}s \right).
\end{equation}
In a LGCP the log intensity $\log x(s) = \eta(s)$ is modelled using a Gaussian process prior $\eta(s) \sim \mathcal{GP}(\mu(s), k(s, s'))$.
@johnson2019spatially obtain Equation \ref{eq:ik2} by considering a discrete Poisson log-linear mixed model approximation to a continuous LGCP, whereby $\eta(s)$ is approximated by a piecewise constant $\eta_i = \mu_i + \phi_i$ in each area $A_i$^[@johnson2019spatially use points generated from a class of inhibition processes which combine sequential inhibition with rejection sampling to evaluate the integral.].
The $i$th discrete spatial random effect is then $\phi_i = \int_{A_i} w_i(s) \phi(s) \text{d}s$,
with covariance structure
\begin{equation}
\text{Cov} \left( \int_{A_i} w_i(s) \phi(s) \text{d}s, \int_{A_j} w_j(s') \phi(s') \text{d}s' \right)
= \int_{A_i} \int_{A_j} w_i(s) w_j(s') k(s, s') \text{d}s\text{d}s',
\end{equation}
corresponding to an areal integrated kernel with a logarithmic link function and Poisson likelihood.

Disaggregation regression (also known as downscaling or interpolation) is another closely related approach which, rather than using the aggregate nature of areal observations as primarily a route towards better area-level estimates, as is our focus, aims to producing high-resolution or point-level estimates from areal observations [@utazi2019spatial; @nandi2020disaggregation].
<!-- Accurate point-level estimates are close to being a necessary intermediate goal towards obtaining accurate area-level estimates. -->
<!-- Disaggregation is challenging without auxiliary information. -->

A benefit of the integration approach is that, if available, additional information accounting for heterogeneity over $A_i$ may be incorperated into the model.
This can be accomplished using weighting distributions $\{W_i\}$ which represent an unequal contribution of each point to the similarity measure, to give a weighted integrated kernel
\begin{equation}
K(A_i, A_j) = \frac{1}{|A_i||A_j|}\int_{A_i} \int_{A_j} w_i(s) w_j(s') k(s, s') \text{d} s \text{d} s', \label{eq:ik2}
\end{equation}
This may be useful in disease mapping, where we expect regions with populations who live close to their shared border to be more strongly correlated than regions whose populations live far apart, which could be accounted for by weighting according to a high resolution measure of population density.

@@

We tested the ability of inferential models with varying spatial random effect specifications to accurately recover small-area quantities.
The data and modelling choices were designed with a spatial epidemiology setting in mind, and are similar to those used in Section \ref{sec:hiv} where we considered models for HIV prevalence using data from national household surveys in sub-Saharan Africa.
The \textsf{R} [@R2021] code used is available from `github.com/athowes/areal-comparison`.
We used the `orderly` package [@orderly] for reproducible research.

@@

### Kernel model details

>>>>

---
title: Understanding models for spatial structure in small-area estimation
author:
  - name: Adam Howes
    email: ath19@ic.ac.uk
    affiliation: A,B
    corresponding: ath19@ic.ac.uk
  - name: Jeffrey W. Eaton
    affiliation: C,B
  - name: Seth R. Flaxman
    affiliation: D
address:
  - code: A
    address: Department of Mathematics, Imperial College London
  - code: B
    address: MRC Centre for Global Infectious Disease Analysis, Imperial College London
  - code: C
    address: Center for Communicable Disease Dynamics, Harvard T.H. Chan School of Public Health, Harvard University
  - code: D
    address: Department of Computer Science, University of Oxford
abstract: |
  |  Spatial correlation between small-areas is typically accounted for within a generalised linear mixed model setting using spatial random effects. However, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic and unsatisfying spatial assumptions when the geometry is not a grid. Furthermore, they do not take into account that across many settings areal data is generated by aggregating continuous point data. Defining spatial random effects which instead correspond to aggregated Gaussian processes gives a more intuitively convincing correlation structure, and allows tools from kernel methods to be used in models for areal data.
  |
  |  In a simulation study, we used a proper scoring rule to evaluate the performance of different spatial random effect specifications. The simulations were performed under model misspecification and across a range of vignette and realistic geometries with different regularities. We also compared model performance in estimating district-level HIV prevalence from PHIA household survey data across four countries in sub-Saharan Africa, using cross-validation and spatial cross-validation.

@@

bibliography: citations.bib
classoption: preprint, authoryear
output:
  rticles::elsevier_article:
    number_sections: true
    includes:
      in_header: preamble.tex
csl: elsarticle.cls
---

@@

Spatial random effects are frequently used to model spatial variation in areal data [@haining2003spatial; @cramb2018investigation].
Gaussian Markov random fields (GMRFs) [@rue2005gaussian], which combine a Gaussian distribution with Markov conditional independence assumptions between areas, are the most common class of models used to specify spatial random effects.
Under these Markov assumptions, observations made in areas close together are correlated, and more distant relationships are ignored.
Arguably the simplest GMRF model is that of @besag1991bayesian, where information is borrowed equally from each adjacent area.
This model is attractive as it requires minimal additional choices and is widely implemented.
As such, it has seen widespread use, including in agriculture [@oliver2015soil], ecology [@saracco2010modeling], epidemiology [@lawson2013statistical], image analysis [@schmid2006bayesian], neuroscience [@gossl2001bayesian] and public health [@dwyer2015drinking].
However, for irregular geometries like the administrative divisions of a country, the assumptions made by the Besag model about space are unrealistic.
As such, we set out to test the hypothesis that using more realistic assumptions about spatial structure would improve the performance of small-area estimation models.
In doing so, we provide practical recommendations and further understanding of models for spatial structure.
<!-- @gutreuter2019improving use the SAR model, which is still adjacency based but is not exactly the Besag model--restructure introduction to be more general and include introduction of other adjacency based models (which all have the same criticisms)? -->

Our motivating application is the small-area estimation of HIV epidemic indicators in sub-Saharan Africa.
Estimates are required to help plan, implement, and evaluate the success of programmes, ensuring that available resources are most effectively used to respond to the epidemic [@hallett2016evaluation; @cuadros2017mapping].
Household surveys provide nationally represenative data about the general population, but are expensive to carry out and as such have small sample sizes at a district level.
Although auxiliary covariates can be used to aid estimation, data on the covariates most strongly associated with HIV, such as sexual risk behaviour [@howes2023spatio] or the prevalence of male circumcision, often face analogous measurement difficulties.
Possibly as a result, models including covariates have only been found to modestly improve predictions \citep[Supplementary Figure 20]{dwyer2019mapping}.
This stands in contrast to other infectious diseases like Malaria where transmission is driven by more predictive and easily-measurable environmental factors [@weiss2015re].
These circumstances foreground the importance of models for spatial structure in HIV mapping.

Within the GRMF framework, attempts to more accurately reflect irregular spatial structure have defined weights specifying the extent to which neighbours are related.
@duncan2017spatial compared seventeen methods for specifying these weights and did not find any which outperform the Besag model, though this conclusion was specific to the often-analysed Scottish lip cancer example, and based on the deviance information criteria (DIC) which is recommended against by @vehtari2017practical.
Another approach is to take into account that areal data is typically produced by aggregating higher resolution data.
Obtaining inference about a variable at a different resolution to that it was observed at is known in geostatistics as the change-of-support problem [@gelfand2001change], and dates to foundational work on block kriging [@krige1951statistical; @matheron1976forecasting].
Ideas from change-of-support modelling are applied by @kelsall2002modeling to consider areal data as coming about by an aggregation of a continuously-indexed Gaussian process, resulting in a covariance structure between two areas given by the average covariance between two points chosen randomly from those areas.
This type of model is particularly natural in the log-Gaussian Cox process modelling framework [@li2012log; @diggle2013spatial; @taylor2015bayesian; @johnson2019spatially].
<!-- `lgcp` [@taylor2015bayesian] and `SDALGCP` [@johnson2019spatially] packages. -->
Inference for aggregated data has recently been advanced by @wilson2018pointless who consider the SPDE approach of @lindgren2011explicit, using `R-INLA`, and an empirical Bayes approach, using `TMB` [@kristensen2015tmb], and made available as a part of the `disaggregation` package [@nandi2020disaggregation].

@@

\newpage

@@

Let $\Sc \subset \mathbb{R}^2$ be the study region, and the disjoint areas $\{A_i\}_{i = 1}^n$ be a spatial partition of $\Sc$.
Areal data are a type of spatial data where observations $y_i$ are associated to areas $A_i$.
Examples include the colour of a pixel, the minimum wage in a state, and the number of disease cases in a region.
Point data are another type of spatial data where instead observations $y(s)$ can be made at any location $s \in \Sc$ of a spatially-continuous stochastic process.
Areal data is often aptly conceptualised as arising from aggregation of point data, such that $y_i = \int_{A_i} y(s) \text{d}s$.
Exceptions include policies determined at an administrative level, such as the minimum wage or disease control measures, which have a genuinely discrete spatial structure.
<!-- Note that, for data of this sort, there is less clear justification for using spatial smoothing. -->
<!-- Realistically, data relate to many covariates, some number of which are likely have discrete spatial structure and others of which continuous spatial structure, such that the result is often an intermediate. -->
<!-- It may be possible to model this explicitly, though the result might look similar to what is achieved using e.g. BYM2. -->

If the areal data we observe indeed arise from aggregation, we may not use the underlying point data because (a) it is unobserved, (b) it may be inaccessible due to privacy constraints, administrative practicality or storage capacity, and (c) the point data are available, but we decide not to use them in the model.
In the final case (c), researchers may prefer to use area-level models as models for point data are typically more complex, require more data, are less immediately applicable to area-level policy making, and have higher computational burden.
Furthermore, seemly point data may actual be areal, such as data observed at polling stations or health facilities which individuals from the surrounding area travel to.
As such, choosing to model the underlying data generating mechanism as either areal or point is often a matter of pragmatism.

@@

Bayesian hierarchical models are an attractive framework for small-area estimation, whereby areal data $\y = (y_1, \ldots, y_n)$ corresponding to the areas $\{A_i\}_{i = 1}^n$, are modelled using a three-layer structure [@berliner1996hierarchical; @cressie2015statistics; @rao2015small] given by
\begin{alignat}{2}
&\text{(Observations)}  &        y_i &\sim p(y_i \, | \, \x, \btheta), \quad i = 1, \ldots, n, \label{eq:data} \\
&\text{(Latent field)}  &        \x &\sim p(\x \, | \, \btheta), \label{eq:process} \\
&\text{(Parameters)}    & \qquad \btheta &\sim p(\btheta), \label{eq:parameters}
\end{alignat}
where $\x$ is a $N$-dimensional latent field, $\btheta$ are the $m$-dimensional parameters, $m < N$.
Spatial structure is modelled at the level of the latent field, and the observation model, typically either binomial or Poisson, is conditionally independent and identically distributed $p(y \, | \, x, \theta) = \prod_{i = 1}^n p(y_i \, | \, x_i, \theta)$.
As small-area estimation models are descriptive rather than mechanistic, it is natural [@best2005comparison] to use a flexible and computationally efficient Gaussian distribution $\x \sim \mathcal{N}(\bmu, \bSigma)$ for the latent field.
Three-layer models with Gaussian latent fields comprise a wide array of popular models, including generalised linear mixed models (GLMMs), and have been collectively studied under the title latent Gaussian models (LGMs) [@rue2009approximate].

@@

Spatial structure can be encoded using a symmetric relation between areas, where $i \sim j$ if the areas $A_i$ and $A_j$ are adjacent or neighbouring.
Adjacency is often defined by a shared border, though other choices, such as inclusion of second or higher-degree neighbours [@paciorek2013spatial], are also possible.
The Besag model [@besag1991bayesian] is a type of improper conditional auto-regressive (ICAR) model where the full conditional distribution of the $i$th spatial random effect is given by
\begin{equation}
    \phi_i \, | \, \phi_{-i} \sim \N \left(\frac{1}{n_{\delta i}} \sum_{j: j \sim i} \phi_j, \frac{1}{n_{\delta i}\tau_\phi}\right), \label{eq:besag}
\end{equation}
where $\delta i$ is the set of neighbours of $A_i$ with cardinality $n_{\delta i} = |\delta i|$ and $\phi_{-i} = (\phi_1, \ldots, \phi_{i - 1}, \phi_{i + 1}, \ldots, \phi_n)^\top$ is the vector of spatial random effects with the $i$th entry removed.
The conditional mean of the random effect $\phi_i$ is the average of its neighbours $\{\phi_j\}_{j \sim i}$ and the precision $n_{\delta i}\tau_\phi$ is proportional to the number of neighbours $n_{\delta i}$.
By Brook's lemma \cite[Chapter 2]{rue2005gaussian} the set of full conditionals of the Besag model are equivalent to the Gaussian Markov random field (GMRF) given by
\begin{equation}
    \phi \sim \mathcal{N}(0, \tau_\phi^{-1} R^{-}), \label{eq:gmrf}
\end{equation}
where $R^{-}$ is the generalised inverse of the rank-deficient structure matrix $R$, so-called because it defines the \textit{structure} of the precision matrix, with entries
\begin{equation}
    R_{ij} =
    \begin{cases}
        n_{\delta i}, & i = j \\
        -1, & i \sim j \\
        0, & \text{otherwise.}
    \end{cases}
\end{equation}
The Markov property arises due to the conditional independence structure $p(\phi_i \, | \, \phi_{-i}) = p(\phi_i \, | \, \phi_{\delta i})$ whereby each area only depends on its neighbours.
This is reflected in the sparsity of $R$ whereby $\phi_i \perp \phi_j \, | \, \phi_{-ij}$ if and only if $R_{ij} = 0$.
The structure matrix $R$ may also be expressed as the Laplacian of the adjacency graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ with vertices $v \in \mathcal{V}$ corresponding to each area and edges $e \in \mathcal{E}$ between vertices $i$ and $j$ when $i \sim j$.
Figure \ref{fig:geometry-graph-zwe} shows the adjacency graph for the districts of Zimbabwe alongside the geometry.
Shortly, in Section \ref{sec:concerns} we discuss the appropriateness of using the graph rather than the complete geometry.

@@

In the previous section we reviewed ways to construct spatial random effect precision matrices using an adjacency relation.
Another approach is to define the covariance matrix using an areal kernel function which gives a measure of similarity between two areas $K: \mathcal{P}(\Sc) \times \mathcal{P}(\Sc) \to \mathbb{R}$, where $\mathcal{P}$ denotes the power set such that $\mathcal{P}(\Sc)$ is the space of subsets of the study region.
If $K$ is positive semi-definite, then we define areal kernel spatial random effects by
\begin{equation}
    \phi \sim \mathcal{N} \left( 0, \frac{1}{\tau_\phi} \mathbf{K} \right), \label{eq:arealkernel}
\end{equation}
where the $n \times n$ Gram matrix $\mathbf{K}$ with entries $K_{ij} = K(A_i, A_j)$ is a valid covariance matrix.
We place $\tau_\phi$ outside of the Gram matrix, analogous to the relation of the precision and structure matrices.
Most well-known spatial process models define the correlation structure between a pair of points using a kernel $k: \mathcal{S} \times \mathcal{S} \to \mathbb{R}$.
We consider a simple method to construct $K$ from $k$: namely averaging the kernel $k$ computed on some collection of points from within each area.

@@

The simplest approach is to use a single point.
A natural choice is the centroid $c_i \in A_i$, given by the arithmetic mean of the latitude and longitude, which we might hope to be representative of the area.
This results in the centroid kernel
\begin{equation}
    K(A_i, A_j) = k(c_i, c_j).
\end{equation}
<!-- More broadly, given a weighting distribution $W_i$ with probability density proportional to $w_i(s)$ over $A_i$, weighted centroids could be used in the above equation. -->
The centroid kernel has been used in an environmental epidemiology setting [@wakefield1999spatial] and in model comparison study [@best2005comparison], both choosing the exponential kernel $k(s_i, s_j) = \exp \left( - |s_i - s_j| / l \right)$; as well as more recently by @teh2021efficient as a part of a spatio-temporal Gaussian process model to infer the reproduction number of COVID-19.
@wakefield1999spatial fix the length-scale $l$ using a semi-variogram and @best2005comparison place a prior on the inverse length-scale.
\citet[Section 3]{best2005comparison} simulated data representing heterogeneous exposure to air pollution, including elevated rates of exposure near two hypothetical point source locations, and found that the centroid kernel tended to over-smooth the high-risk areas, though it is unsuprising that a stationary covariance would struggle to recover non-stationary structure.

@@

Rather than looking to choose a single representative point, an alternative is to represent the whole area by integrating the kernel over the areas of interest.
This results in the integrated kernel
\begin{equation}
K(A_i, A_j) = \frac{1}{|A_i||A_j|} \int_{A_i} \int_{A_j} k(s, s') \text{d} s \text{d} s'. \label{eq:ik1}
\end{equation}
This results in equivalent covariance structure to that obtained by aggregating a spatially continuous Gaussian process with kernel $k$ over the areal partition, studied in the machine learning literature under the name aggregated Gaussian processes [@law2018variational; @tanaka2019spatially; @yousefi2019multi; @hamelijnck2019multi].å
<!-- @law2018variational develop a variational Bayesian inference method for aggregated observation models. -->
<!-- @tanaka2019spatially consider aggregation over a linear combination of latent Gaussian processes, each of which is associated to data corresponding to a different partition of the study region. -->
Unlike for the centroid kernel where $K_{ii} = 1$ for all $i$, the marginal variance of the $i$th spatial random effect $K_{ii} = K(A_i, A_i)$ varies depending on the area: becoming smaller for more compact areas and larger for areas which are of greater extent or more spread out.

If available, additional information accounting for heterogeneity over $A_i$ may be incorporated into the integrated kernel.
This can be accomplished using weighting distributions $\{W_i\}$ which represent an unequal contribution of each point to the similarity measure, to give a weighted integrated kernel
\begin{equation}
K(A_i, A_j) = \frac{1}{|A_i||A_j|}\int_{A_i} \int_{A_j} w_i(s) w_j(s') k(s, s') \text{d} s \text{d} s', \label{eq:ik2}
\end{equation}
This may be useful in disease mapping, where we expect regions with populations who live close to their shared border to be more strongly correlated than regions whose populations live far apart, which could be accounted for by weighting according to a high resolution measure of population density.

@@

We tested the ability of inferential models with varying spatial random effect specifications to accurately recover small-area quantities.
The data and modelling choices were designed with a spatial epidemiology setting in mind.
The \textsf{R} [@R2021] code used is available from `github.com/athowes/areal-comparison`.
We used the `orderly` package [@orderly] for reproducible research.

@@

### Kernels
